Deep mutational scanning of hemagglutinin helps distinguish the evolutionary fate of human H3N2 influenza virus lineages

### 2018-01-17 paper/manuscript.tex
Some additions of detail to methods

\matmethods{Deep sequencing data are available from the Sequence Read Archive under BioSample accessions SAMN08102609 and SAMN08102610. 
Computer code used to analyze the data and produce the results in the paper are in...

@@

To generate wildtype Perth/2009 H3N2 virus, we transfected a co-culture of 293T and MDCK-SIAT1-TMPRSS2 cells with reverse genetics plasmids encoding the Perth/2009 HA and NA, WSN/1933 internal genes, and the TMPRSS2 lentiviral vector.
\comment{add more details about reverse genetics?}
The viral supernatant was harvested at 72 hours post-transfection, clarified at 1200 rpm for 5 min, and 0.01 to 1.0 ul of the supernatant was passaged onto $4 \times 10^{5}$ MDCK-SIAT1-TMPRSS2 cells in six-well plates for a total of six serial passages.
After the final passage, RNA was extracted from 140 ul of the viral supernatant using the QIAamp Viral RNA Mini Kit.
The HA and NA genes were reverse transcribed, PCR amplified, and sequenced by Sanger sequencing.
Two mutations, G78D and T212I, arose to fixation after serial passaging. 
The Perth/2009 HA carrying these two mutations were cloned into the pHW2000~\citep{hoffmann2000dna} and pICR2~\cite{ashenberg2017deep} plasmid backbones.

@@

The codon-mutant libraries were generated in the Perth/2009 HA-G78D-T212I background using a PCR-based approach described in~\cite{bloom2014experimentally,dingens2017comprehensive} using two rounds of mutagenesis.
The codon mutations were introduced into three independent replicates of the Perth/2009 HA variant.
The mutant variants were then cloned into the pICR2 vector by digestion with BsmBI restriction sites, ligation by T4 DNA ligase, and electroporation and transformation into ElectroMAX DH10B competent cells (Invitrogen; 18290015). 
We pooled over 6 million transformants for each replicate, cultured in liquid LB + ampicillin at 37$^{\circ}$C for 3 h with shaking, and maxiprepped.
We randomly chose 31 clones to Sanger sequence, and summary statistics from sequencing are shown in Figure~\ref{suppfig:SangerSeq}.

@@

To generate the mutant virus libraries, we transfected $5 \times 10^5$ MDCK-SIAT1-TMPRSS2 cells in suspension with 937.5 ng each of four protein expression plasmids encoding the ribonucleoprotein complex (HDM-Nan95-PA, HDM-Nan95-PB1, HDM-Nan95-PB2, and HDM-Aichi68-NP)~\cite{gong2013stability}, and 1250 ng of one of the three pICR2-mutant-HA libraries (or the wildtype pICR2-Perth2009-HA-G78D-T212I control) using Lipofectamine3000 (ThermoFisher;L3000008).
We allowed the transfected cells to adhere in 6-well plates, then four hours later changed the media to D10 media (DMEM supplemented with 10\% heat-inactivated FBS, 2 mM L-glutamine, 100 U of penicillin/mL, and 100 $\mu$g of streptomycin/mL).
Eighteen hours after transfection, we infected the cells with WSN/1933 HA-deficient helper virus by preparing an inoculum of 500 TCID$_{50}$ per $\mu$L in influenza growth media (Opti-MEM supplemented with 0.01\% heat-inactivated FBS, 0.3\% BSA, 100 U of penicillin/mL, 100 $\mu$g of streptomycin/mL, and 100 $\mu$g of calcium chloride/mL), and aspirating the D10 media from the cells, and adding 2 mL of the helper-virus inoculum to each well.
After three hours, we changed the media to fresh influenza growth media.
Twenty-four hours after helper-virus infection, we harvested the viral supernatants for each replicate, froze aliquots at -80$^{\circ}$C, and titered in MDCK-SIAT1-TMPRSS2 cells.

@@

\subsection*{Alignments and phylogenetic analysis of HA sequences}
To build the ExpCM using our measurements, we first downloaded all full-length H3 HA sequences from the Influenza Virus Resource~\cite{bao2008}, and randomly subsampled two sequences per year.
The subsampled sequences were then aligned using MAFFT~\cite{katoh2013mafft} and used to build a phylogenetic tree using RAxML~\cite{stamatakis2006raxml}.
We then used the phydms program~\cite{hilton2017phydms} (\url{https://github.com/jbloomlab/phydms})

@@

\begin{suppfigure}
\centerline{\includegraphics[width=0.6\textwidth]{figs/S04_seqpref_zoomed/seqpref_zoomed.pdf}}
\caption{\label{suppfig:seqpref_zoomed}
{\bf The per-site sequence preference at epitope and non-epitope sites}
The per-site sequence preferences shown in Figure~\ref{fig:sequence_preference}B, but on their own y-axes.
}
\end{suppfigure}

>>>>

\matmethods{Deep sequencing data are available from the Sequence Read Archive under BioSample accessions SAMN08102609 and SAMN08102610. 
Computer code used to analyze the data and produce the results in the paper are in \comment{link to repository}.

@@

To generate wildtype Perth/2009 H3N2 virus, we transfected in triplicate co-cultures of 293T and MDCK-SIAT1-TMPRSS2 cells with reverse genetics plasmids encoding the Perth/2009 HA and NA, WSN/1933 internal genes, and the TMPRSS2 lentiviral vector.
\comment{add more details about reverse genetics?}
The viral supernatant was harvested at 72 hours post-transfection, clarified at 1200 rpm for 5 min, and 0.01 to 1.0 ul of the supernatant was passaged onto $4 \times 10^{5}$ MDCK-SIAT1-TMPRSS2 cells in six-well plates for a total of six serial passages.
After the final passage, RNA was extracted from 140 ul of the viral supernatant for each replicate using the QIAamp Viral RNA Mini Kit.
The HA and NA genes were reverse transcribed, PCR amplified, and sequenced by Sanger sequencing.
Two mutations, G78D and T212I, arose to fixation in one of the replicates after serial passaging. 
The Perth/2009 HA carrying these two mutations were cloned into the pHW2000~\citep{hoffmann2000dna} and pICR2~\cite{ashenberg2017deep} plasmid backbones.

@@

The codon-mutant libraries were generated in the Perth/2009 HA-G78D-T212I background using a PCR-based approach described in~\cite{bloom2014experimentally,dingens2017comprehensive} using two rounds of mutagenesis.
The script used to design the mutagenesis primers is on \url{https://github.com/jbloomlab/CodonTilingPrimers}.
The codon mutations were introduced into three independent replicates of the Perth/2009 HA variant.
The mutant variants were then cloned into the pICR2 vector by digestion with BsmBI restriction sites, ligation by T4 DNA ligase, and electroporation and transformation into ElectroMAX DH10B competent cells (Invitrogen; 18290015). 
We pooled over 6 million transformants for each replicate, cultured in liquid LB + ampicillin at 37$^{\circ}$C for 3 h with shaking, and maxiprepped.
We randomly chose 31 clones to Sanger sequence, and summary statistics from sequencing are shown in Figure~\ref{suppfig:SangerSeq}.

@@

To generate the mutant virus libraries, we transfected $5 \times 10^5$ MDCK-SIAT1-TMPRSS2 cells in suspension with 937.5 ng each of four protein expression plasmids encoding the ribonucleoprotein complex (HDM-Nan95-PA, HDM-Nan95-PB1, HDM-Nan95-PB2, and HDM-Aichi68-NP)~\cite{gong2013stability}, and 1250 ng of one of the three pICR2-mutant-HA libraries (or the wildtype pICR2-Perth2009-HA-G78D-T212I control) using Lipofectamine3000 (ThermoFisher;L3000008).
We allowed the transfected cells to adhere in 6-well plates, then four hours later changed the media to D10 media (DMEM supplemented with 10\% heat-inactivated FBS, 2 mM L-glutamine, 100 U of penicillin/mL, and 100 $\mu$g of streptomycin/mL).
Eighteen hours after transfection, we infected the cells with WSN/1933 HA-deficient helper virus by preparing an inoculum of 500 TCID$_{50}$ per $\mu$L in influenza growth media (Opti-MEM supplemented with 0.01\% heat-inactivated FBS, 0.3\% BSA, 100 U of penicillin/mL, 100 $\mu$g of streptomycin/mL, and 100 $\mu$g of calcium chloride/mL), aspirating the D10 media from the cells, and adding 2 mL of the helper-virus inoculum to each well.
After three hours, we changed the media to fresh influenza growth media.
Twenty-four hours after helper-virus infection, we harvested the viral supernatants for each replicate, froze aliquots at -80$^{\circ}$C, and titered in MDCK-SIAT1-TMPRSS2 cells.

@@

\subsection*{Alignments and phylogenetic analysis of HA sequences}
To build the ExpCM using our measurements, we first downloaded all full-length H3 HA sequences from the Influenza Virus Resource~\cite{bao2008}, and randomly subsampled two sequences per year.
The subsampled sequences were then aligned using MAFFT~\cite{katoh2013mafft} and used to build a phylogenetic tree using RAxML~\cite{stamatakis2006raxml}.
We then used the phydms program~\cite{hilton2017phydms} (\url{https://github.com/jbloomlab/phydms}) to fit phylogenetic substitution models using the aligned subsampled sequences.

@@

\begin{suppfigure}
\centerline{\includegraphics[width=0.6\textwidth]{figs/S04_seqpref_zoomed/seqpref_zoomed.pdf}}
\caption{\label{suppfig:seqpref_zoomed}
{\bf The per-site sequence preference at epitope and non-epitope sites.}
The per-site sequence preferences shown in Figure~\ref{fig:sequence_preference}B, but on separate y-axes.
}
\end{suppfigure}
